---
title: "final_result"
author: "Abdelrehim Sabri (500998232)"
date: '2019-11-25'
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install R packages

```{r install_rmarkdown, eval=FALSE}
#install.packages("rpart")
#install.packages("caret")
#install.packages("e1071")
#install.packages("randomForest")
#install.packages("corrplot")
#install.packages("caretEnsemble")
#install.packages("lsr")
```

## Exploratory Data Analysis

```{r }
options(warm=-1)
library(rsample)  # data splitting
library(lubridate)
library(ggplot2)  # data visualization
library(randomForest)
library(dplyr) # data transformation
#library(rpart)
library(caret) # implementing with caret
library(caretEnsemble)
library(e1071)
library(corrplot)
library(unbalanced)
library(mlbench)
library(pROC)
library(lsr)
```

## Includes functions to clean datasets

Read datasets from csv file
```{r }
build_clean_dataset <- function() {
  datasetloc = "C:/Projects/temp/clean/Health_Care_History.csv"
  if (file.exists(datasetloc)) {
    alldata <- read.csv(file=datasetloc, header = T)
  }
  return(alldata)
}
```

Convert the date to age and group them into four groups (0-25, 26-40, 41-50, 50-65, 65+)
```{r }
age <- function(dob, age.day = today(), units = "years", floor = TRUE) {
  calc.age = interval(dob, age.day) / duration(num = 1, units = units)
  if (floor) return(as.integer(floor(calc.age)))
  
  return(calc.age)
}

get_age_group <- function(a) {
  ifelse(a<25,25, ifelse(a<40, 40, ifelse(a<50,50,65)))
}
```

Group the countries of the patients based on ethnic groups
```{r }
east_europe <- c('Ukraine','Russia','Poland','Czech Republic','Hungary')
west_europe <- c('Austria','Belgium','France','Germany','Italy','Netherlands','Portugal','Spain','Switzerland')
north_europe <- c('Sweden', 'Finland', 'Denmark')
british <- c('England','Scotland','Ireland')

get_ethnic_group <- function(country) {
  ifelse((country %in% east_europe), 'east_europe',
         ifelse((country %in% west_europe) ,'west_europe',
                ifelse((country %in% north_europe), 'north_europe',
                       ifelse((country %in% british), 'british',
                              country))))
}
```

Read the dataset and remove patient ids from the analysis 
```{r}
patients <- build_clean_dataset()

#remove the patient ids from the dataset
patients <- patients[,-1]
str(patients)
```

```{r}
summary(patients)
```
From the summary, there are no NaN nor missing data in the dataset.

Fix the education column values by fixing the misspelled words
```{r}
patients$education <- ifelse(patients$education == 'highscool', as.character('highschool'), as.character(patients$education))
patients$education <- ifelse(as.factor(patients$education) == 'phD/MD', as.character('phd/md'), as.character(patients$education))
patients$education <- as.factor(patients$education)
```

Group the ancestry countries to ethnic groups
```{r}
patients$ancestry <- as.factor(get_ethnic_group(patients$ancestry))
```

Convert the date of birth into age and group them into 25 40 50 65
```{r}
patients$age <- age(patients$dob)
patients$age <- get_age_group(age(patients$dob))
```

For the analysis purposes, move each disease to separate column with binary values, where 0: patient does not has the disease and 1: patient has the disease 
```{r}
get_binary_value <- function(value, compare_to) {
  ifelse(value==compare_to,1,0)
}
patients$prostate_cancer <- get_binary_value(patients$disease,'prostate cancer')
patients$skin_cancer <- get_binary_value(patients$disease,'skin cancer')
patients$breast_cancer <- get_binary_value(patients$disease,'breast cancer')
patients$hiv_aids <- get_binary_value(patients$disease,'HIV/AIDS')
patients$diabetes <- get_binary_value(patients$disease,'diabetes')
patients$heart_disease <- get_binary_value(patients$disease,'heart disease')
patients$hypertension <- get_binary_value(patients$disease,'hypertension')
patients$endometriosis <- get_binary_value(patients$disease,'endometriosis')
patients$multiple_sclerosis <- get_binary_value(patients$disease,'multiple sclerosis')
patients$schizophrenia <- get_binary_value(patients$disease,'schizophrenia')
patients$kidney_disease <- get_binary_value(patients$disease,'kidney disease')
patients$gastritis <- get_binary_value(patients$disease,'gastritis')
patients$alzheimer <- get_binary_value(patients$disease,'Alzheimer disease')
str(patients)
```

## Use barplots for the distribution of the categorical columns

Draw a bar plot to count the total number of diseases in the dataset
```{r}
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin.

disease_counts <- table(patients$disease)
barplot(sort(disease_counts, decreasing = TRUE), main="Disease Names", 
        xlab="Diseases Frequency", 
        col=rainbow(20),
        horiz=TRUE,
        cex.names=0.8,
        xlim = c(0, 350))
```

Observation : Male are more sick than Female
```{r}
gender_counts <- table(patients$gender)
barplot(sort(gender_counts, decreasing = TRUE), main="Gender", 
        col=rainbow(20), las=1)
```

Observation : age group that are more sick
```{r}
age_breaks <- c(0,25,40,65,100)
tags <- c("[0-25)","[26-40)", "[41-65)", "[65+)")
age_group_tags <- cut(patients$age, 
                  breaks=age_breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
summary(age_group_tags)
#age_counts <- table(patients$age)
age_counts <- table(age_group_tags)
barplot(sort(age_counts, decreasing = TRUE), main="Age",
        col=rainbow(20), las=1)
```

Observation : Disease and Gender distrubution
```{r}
disease_name = c(as.character(unique(patients$disease)))

for (d in disease_name) {
  gender_disease_counts <- subset(patients, patients$disease == d)
  gender_disease_counts <- table(gender_disease_counts$gender)
  barplot(gender_disease_counts, main=d, col=rainbow(20), las=1)
}
```

Observation : Disease and ancestry distrubution
```{r}
for (d in disease_name) {
  ancestry_disease_counts <- subset(patients, patients$disease == d)
  ancestry_disease_counts <- table(ancestry_disease_counts$ancestry)
  barplot(ancestry_disease_counts, main=d, col=rainbow(20), las=1)
}
```
## Plots for dependent variables

Observation : Disease and age distrubution
```{r}
for (d in disease_name) {
  age_disease_counts <- subset(patients, patients$disease == d)
  #age_disease_counts <- table(age_disease_counts$age_group_tags)
  age_disease_counts <- table(age_disease_counts$age)
  barplot(age_counts, main=d, col=rainbow(20), las=1)
}
```


Observation : Disease and employment status distrubution
```{r}
for (d in disease_name) {
  emp_disease_counts <- subset(patients, patients$disease == d)
  emp_disease_counts <- table(emp_disease_counts$employment_status)
  barplot(emp_disease_counts, main=d, col=rainbow(20), las=1)
}
```
Observation : Disease and number of children
```{r}
for (d in disease_name) {
  child_disease_counts <- subset(patients, patients$disease == d)
  child_disease_counts <- table(child_disease_counts$children)
  barplot(child_disease_counts, main=d, col=rainbow(20), las=1)
}
```
Observation : Disease and avg commute
```{r}
for (d in disease_name) {
  comm_disease_counts <- subset(patients, patients$disease == d)
  comm_disease_counts <- table(comm_disease_counts$avg_commute)
  barplot(comm_disease_counts, main=d, col=rainbow(20), las=1)
}
```
Observation : Disease and daily internet use
```{r}
for (d in disease_name) {
  net_disease_counts <- subset(patients, patients$disease == d)
  net_disease_counts <- table(net_disease_counts$daily_internet_use)
  barplot(net_disease_counts, main=d, col=rainbow(20), las=1)
}
```
Observation : Disease and available vehicles
```{r}
for (d in disease_name) {
  veh_disease_counts <- subset(patients, patients$disease == d)
  veh_disease_counts <- table(veh_disease_counts$available_vehicles)
  barplot(veh_disease_counts, main=d, col=rainbow(20), las=1)
}
```
Observation : Disease and military service
```{r}
for (d in disease_name) {
  mil_disease_counts <- subset(patients, patients$disease == d)
  mil_disease_counts <- table(mil_disease_counts$military_service)
  barplot(mil_disease_counts, main=d, col=rainbow(20), las=1)
}
```

## Feature Selection

Will do feature selection using two methods, Chi-squared and Cramer's V after splitting and balancing the dataset for three diseases (alzheimer, hypertension, skin cancer)

Feature the selection Alsheimer using Chi-squared 
```{r}
alzheimer_set <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service, alzheimer)
FeatureTrain <- sample(nrow(alzheimer_set), 0.7*nrow(alzheimer_set), replace = FALSE)
FeatureTrainSet <- alzheimer_set[FeatureTrain,]
FeatureValidSet <- alzheimer_set[-FeatureTrain,]

response <- as.factor(patients$alzheimer)
input <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service)

data <- ubOver(X=input, Y=response)
alzheime_os_dataset <- cbind(data$X, class=data$Y)

chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$gender)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$age)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$education)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$marital_status)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$zipcode)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$employment_status)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$children)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$ancestry)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$avg_commute)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$daily_internet_use)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$available_vehicles)
chisq.test(alzheime_os_dataset$class, alzheime_os_dataset$military_service)


alzheime_os_dataset %>%
  filter(class == "1") %>%
  select_if(is.numeric) %>%
  cor() %>%
  corrplot::corrplot()
```

Feature the selection Alsheimer using Cramer's V 
```{r}
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$gender)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$age)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$education)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$marital_status)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$zipcode)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$employment_status)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$children)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$ancestry)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$avg_commute)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$daily_internet_use)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$available_vehicles)
cramersV(alzheime_os_dataset$class, alzheime_os_dataset$military_service)
```

Feature the selection Hypertension using Chi-squared
```{r}
hypertension_set <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service, hypertension)
FeatureTrain <- sample(nrow(hypertension_set), 0.7*nrow(hypertension_set), replace = FALSE)
FeatureTrainSet <- hypertension_set[FeatureTrain,]
FeatureValidSet <- hypertension_set[-FeatureTrain,]

response <- as.factor(patients$hypertension)
input <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service)

data <- ubOver(X=input, Y=response)
hypertension_os_dataset <- cbind(data$X, class=data$Y)

chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$gender)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$age)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$education)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$marital_status)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$zipcode)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$employment_status)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$children)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$ancestry)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$avg_commute)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$daily_internet_use)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$available_vehicles)
chisq.test(hypertension_os_dataset$class, hypertension_os_dataset$military_service)
```

Feature the selection Hypertension using Cramer's V 
```{r}
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$gender)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$age)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$education)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$marital_status)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$zipcode)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$employment_status)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$children)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$ancestry)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$avg_commute)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$daily_internet_use)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$available_vehicles)
cramersV(hypertension_os_dataset$class, hypertension_os_dataset$military_service)
```

Feature the selection Skin Cancer using Chi-squared
```{r}
skin_cancer_set <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service, skin_cancer)
FeatureTrain <- sample(nrow(skin_cancer_set), 0.7*nrow(skin_cancer_set), replace = FALSE)
FeatureTrainSet <- skin_cancer_set[FeatureTrain,]
FeatureValidSet <- skin_cancer_set[-FeatureTrain,]

response <- as.factor(patients$skin_cancer)
input <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service)

data <- ubOver(X=input, Y=response)
skin_cancer_os_dataset <- cbind(data$X, class=data$Y)

chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$gender)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$age)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$education)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$marital_status)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$zipcode)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$employment_status)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$children)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$ancestry)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$avg_commute)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$daily_internet_use)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$available_vehicles)
chisq.test(skin_cancer_os_dataset$class, skin_cancer_os_dataset$military_service)
```

Feature the selection Skin Cancer using Cramer's V
```{r}
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$gender)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$age)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$education)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$marital_status)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$zipcode)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$employment_status)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$children)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$ancestry)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$avg_commute)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$daily_internet_use)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$available_vehicles)
cramersV(skin_cancer_os_dataset$class, skin_cancer_os_dataset$military_service)
```

## Modeling

After having better understanding of the data, will begin preparing for modeling in order to predict diseases.

The process will be divided the process into steps:
1. Dealing with the Imbalance
2. Define algorithms
3. Testing algorithms

## Dealing with the Imbalance

From the exploratory analysis abovem the dependent variable is imbalanced. There are many alternatives to tackle this problem:
* Over-sampling
* Under-sampling
* Synthetic Minority Over-Sampling Technique (SMOTE) Sampling
* Cost Sensitive Learning

For this dataset, will use over-sampling and SMOTE technique.

## Patients with alzheimer
```{r}
#Convert all columns to factor
patients[] <- lapply( patients, factor) # the "[]" keeps the dataframe structure
 col_names <- names(patients)
 patients[col_names] <- lapply(patients[col_names], factor)
 
#See the data before balancing
barplot(table(patients$alzheimer), xlab=colnames(patients$alzheimer))

#filter the dataset and have only alzheimer disease as the target 
alzheimer_set <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles,zipcode, children,military_service, alzheimer)

#The data were partitioned into a test and training set using a 70/30 split.
train <- sample(nrow(alzheimer_set), 0.7*nrow(alzheimer_set), replace = FALSE)
  TrainSet <- alzheimer_set[train,]
  ValidSet <- alzheimer_set[-train,]
  
response <- as.factor(patients$alzheimer)
input <- select(patients, gender, age, employment_status, education, marital_status, ancestry)
```

### Excerice the Undersampling, oversampling, and smote against the test dataset

## Logistic Regression, Randomforest, and Naive Bayes Models

```{r}
  #initialize variables
  us_glm_accuracy <- c()
  us_glm_precision <- c()
  us_glm_recall <- c()
  us_glm_f1 <- c()
  
  os_glm_accuracy <- c()
  os_glm_precision <- c()
  os_glm_recall <- c()
  os_glm_f1 <- c()
  
  smote_glm_accuracy <- c()
  smote_glm_precision <- c()
  smote_glm_recall <- c()
  smote_glm_f1 <- c()
  
  us_rf_accuracy <- c()
  us_rf_precision <- c()
  us_rf_recall <- c()
  us_rf_f1 <- c()
  
  os_rf_accuracy <- c()
  os_rf_precision <- c()
  os_rf_recall <- c()
  os_rf_f1 <- c()
  
  smote_rf_accuracy <- c()
  smote_rf_precision <- c()
  smote_rf_recall <- c()
  smote_rf_f1 <- c()
  
  us_nb_accuracy <- c()
  us_nb_precision <- c()
  us_nb_recall <- c()
  us_nb_f1 <- c()
  
  os_nb_accuracy <- c()
  os_nb_precision <- c()
  os_nb_recall <- c()
  os_nb_f1 <- c()
  
  smote_nb_accuracy <- c()
  smote_nb_precision <- c()
  smote_nb_recall <- c()
  smote_nb_f1 <- c()
  
  #use the 10-fold cross-validation and repeate the step 3 times
  train_control <- trainControl(method = "cv", number = 10)
  metric <- "Accuracy"
  mtry <- sqrt(ncol(alzheimer_set))
  tunegrid <- expand.grid(.mtry=mtry)    
```

```{r}
  # iterate throug the sampling and model 10 times and get the mean to get the best model for the dataset prediction
  for (i in 1:10) {
    
    seed <- 999+i
    set.seed(seed)
    
    #run the undersampling
    data <- ubUnder(X=input, Y=response, perc=40, method="percPos")
    us_dataset <- cbind(data$X, class=data$Y)
    
    #run the oversampling
    data <- ubOver(X=input, Y=response)
    os_dataset <- cbind(data$X, class=data$Y)
    
    #run the smote
    data <- ubSMOTE(X=input, Y=response)
    smote_dataset <- cbind(data$X, class=data$Y)
    
    #run the logistic regression for the undersampling
    glm_mod <- caret::train(class~.,data=us_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    us_glm_accuracy <- c(us_glm_accuracy, us_cm$overall['Accuracy'])
    us_glm_precision <- c(us_glm_precision, us_cm$byClass['Precision'])
    us_glm_recall <- c(us_glm_recall, us_cm$byClass['Recall'])
    us_glm_f1 <- c(us_glm_f1, us_cm$byClass['F1'])
    
    #run the logistic regression for the oversampling
    glm_mod <- caret::train(class~.,data=os_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    os_glm_accuracy <- c(os_glm_accuracy, os_cm$overall['Accuracy'])
    os_glm_precision <- c(os_glm_precision, os_cm$byClass['Precision'])
    os_glm_recall <- c(os_glm_recall, os_cm$byClass['Recall'])
    os_glm_f1 <- c(os_glm_f1, os_cm$byClass['F1'])
    
    #run the logistic regression for the smote
    glm_mod <- caret::train(class~.,data=smote_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    smote_glm_accuracy <- c(smote_glm_accuracy, cm_smote$overall['Accuracy'])
    smote_glm_precision <- c(smote_glm_precision, cm_smote$byClass['Precision'])
    smote_glm_recall <- c(smote_glm_recall, cm_smote$byClass['Recall'])
    smote_glm_f1 <- c(smote_glm_f1, cm_smote$byClass['F1'])

    
    #run the random forest for the undersampling
    rf_mod <- caret::train(class~., data=us_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    us_rf_accuracy <- c(us_rf_accuracy, us_cm$overall['Accuracy'])
    us_rf_precision <- c(us_rf_precision, us_cm$byClass['Precision'])
    us_rf_recall <- c(us_rf_recall, us_cm$byClass['Recall'])
    us_rf_f1 <- c(us_rf_f1, us_cm$byClass['F1'])
    
    #run the random forest for the oversampling
    rf_mod <- caret::train(class~., data=os_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    os_rf_accuracy <- c(os_rf_accuracy, os_cm$overall['Accuracy'])
    os_rf_precision <- c(os_rf_precision, os_cm$byClass['Precision'])
    os_rf_recall <- c(os_rf_recall, os_cm$byClass['Recall'])
    os_rf_f1 <- c(os_rf_f1, os_cm$byClass['F1'])
    
    #run the random forest for the smote
    rf_mod <- caret::train(class~., data=smote_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    smote_rf_accuracy <- c(smote_rf_accuracy, cm_smote$overall['Accuracy'])
    smote_rf_precision <- c(smote_rf_precision, cm_smote$byClass['Precision'])
    smote_rf_recall <- c(smote_rf_recall, cm_smote$byClass['Recall'])
    smote_rf_f1 <- c(smote_rf_f1, cm_smote$byClass['F1'])
    
    #run the naive byes for the undersampling
    nb_mod <- caret::train(class~., data=us_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    us_nb_accuracy <- c(us_nb_accuracy, us_cm$overall['Accuracy'])
    us_nb_precision <- c(us_nb_precision, us_cm$byClass['Precision'])
    us_nb_recall <- c(us_nb_recall, us_cm$byClass['Recall'])
    us_nb_f1 <- c(us_nb_f1, us_cm$byClass['F1'])

    #run the naive byes for the oversampling
    nb_mod <- caret::train(class~., data=os_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    os_nb_accuracy <- c(os_nb_accuracy, os_cm$overall['Accuracy'])
    os_nb_precision <- c(os_nb_precision, os_cm$byClass['Precision'])
    os_nb_recall <- c(os_nb_recall, os_cm$byClass['Recall'])
    os_nb_f1 <- c(os_nb_f1, os_cm$byClass['F1'])
    
    #run the naive byes for the smote
    nb_mod <- caret::train(class~., data=smote_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$alzheimer), mode='everything')
    smote_nb_accuracy <- c(smote_nb_accuracy, cm_smote$overall['Accuracy'])
    smote_nb_precision <- c(smote_nb_precision, cm_smote$byClass['Precision'])
    smote_nb_recall <- c(smote_nb_recall, cm_smote$byClass['Recall'])
    smote_nb_f1 <- c(smote_nb_f1, cm_smote$byClass['F1'])
  }
  
```




##################################################################################################################################


## Result of the alzheimer analysis

The data were partitioned into a test and training set using a 70/30 split.
```{r}
df <- data.frame(us_glm_accuracy, os_glm_accuracy, smote_glm_accuracy, us_rf_accuracy, os_rf_accuracy, smote_rf_accuracy, us_nb_accuracy, os_nb_accuracy, smote_nb_accuracy)

us_glm_accuracy
os_glm_accuracy
smote_glm_accuracy
us_rf_accuracy
os_rf_accuracy
smote_rf_accuracy
us_nb_accuracy
os_nb_accuracy
smote_nb_accuracy

us_glm_precision
os_glm_precision
smote_glm_precision
us_rf_precision
os_rf_precision
smote_rf_precision
us_nb_precision
os_nb_precision
smote_nb_precision

us_glm_recall
os_glm_recall
smote_glm_recall
us_rf_recall
os_rf_recall
smote_rf_recall
us_nb_recall
os_nb_recall
smote_nb_recall

us_glm_f1
os_glm_f1
smote_glm_f1
us_rf_f1
os_rf_f1
smote_rf_f1
us_nb_f1
os_nb_f1
smote_nb_f1

c1 <- rainbow(10)
c2 <- rainbow(10, alpha=0.2)
c3 <- rainbow(10, v=0.7)
boxplot(df, col=c2, medcol=c3, whiskcol=c1, staplecol=c3, boxcol=c3, outcol=c3, pch=23, cex=2)

mean(us_nb_accuracy)
mean(us_nb_precision)
mean(us_nb_recall)
mean(us_nb_f1)

mean(os_nb_accuracy)
mean(os_nb_precision)
mean(os_nb_recall)
mean(os_nb_f1)

mean(smote_nb_accuracy)
mean(smote_nb_precision)
mean(smote_nb_recall)
mean(smote_nb_f1)

a <- matrix(
  c(mean(us_glm_accuracy),mean(us_glm_precision),mean(us_glm_recall),mean(us_glm_f1),
    mean(os_glm_accuracy),mean(os_glm_precision),mean(os_glm_recall),mean(os_glm_f1),
    mean(smote_glm_accuracy),mean(smote_glm_precision),mean(smote_glm_recall),mean(smote_glm_f1)),
  nrow=3,
  ncol=4,
  byrow = TRUE
)

a
```

## Patients with hypertension
```{r}
#See the data before balancing
barplot(table(patients$hypertension), xlab=colnames(patients$hypertension))

#filter the dataset and have only hypertension disease as the target 
hypertension_set <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service, hypertension)

#The data were partitioned into a test and training set using a 70/30 split.
train <- sample(nrow(hypertension_set), 0.7*nrow(hypertension_set), replace = FALSE)
  TrainSet <- hypertension_set[train,]
  ValidSet <- hypertension_set[-train,]
  
response <- as.factor(patients$hypertension)
input <- select(patients, gender, age, employment_status, education, marital_status, ancestry)
```

### Excerice the Undersampling, oversampling, and smote against the test dataset

## Logistic Regression, Randomforest, and Naive Bayes Models
```{r}
  #initialize variables
  us_glm_accuracy <- c()
  us_glm_precision <- c()
  us_glm_recall <- c()
  us_glm_f1 <- c()
  
  os_glm_accuracy <- c()
  os_glm_precision <- c()
  os_glm_recall <- c()
  os_glm_f1 <- c()
  
  smote_glm_accuracy <- c()
  smote_glm_precision <- c()
  smote_glm_recall <- c()
  smote_glm_f1 <- c()
  
  us_rf_accuracy <- c()
  us_rf_precision <- c()
  us_rf_recall <- c()
  us_rf_f1 <- c()
  
  os_rf_accuracy <- c()
  os_rf_precision <- c()
  os_rf_recall <- c()
  os_rf_f1 <- c()
  
  smote_rf_accuracy <- c()
  smote_rf_precision <- c()
  smote_rf_recall <- c()
  smote_rf_f1 <- c()
  
  us_nb_accuracy <- c()
  us_nb_precision <- c()
  us_nb_recall <- c()
  us_nb_f1 <- c()
  
  os_nb_accuracy <- c()
  os_nb_precision <- c()
  os_nb_recall <- c()
  os_nb_f1 <- c()
  
  smote_nb_accuracy <- c()
  smote_nb_precision <- c()
  smote_nb_recall <- c()
  smote_nb_f1 <- c()
  
  #use the 10-fold cross-validation and repeate the step 3 times
  train_control <- trainControl(method = "cv", number = 10)
  metric <- "Accuracy"
  mtry <- sqrt(ncol(alzheimer_set))
  tunegrid <- expand.grid(.mtry=mtry)    
```

```{r}
  # iterate throug the sampling and model 10 times and get the mean to get the best model for the dataset prediction
  for (i in 1:10) {
    
    #run the undersampling
    data <- ubUnder(X=input, Y=response, perc=40, method="percPos")
    us_dataset <- cbind(data$X, class=data$Y)
    
    #run the oversampling
    data <- ubOver(X=input, Y=response)
    os_dataset <- cbind(data$X, class=data$Y)
    
    #run the smote
    data <- ubSMOTE(X=input, Y=response)
    smote_dataset <- cbind(data$X, class=data$Y)
    
    #use the 10-fold cross-validation and repeate the step 3 times
    train_control <- trainControl(method = "repeatedcv", number = 10, repeats=3, savePredictions = TRUE)
    
    #run the logistic regression for the undersampling
    glm_mod <- caret::train(class~.,data=us_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    us_glm_accuracy <- c(us_glm_accuracy, us_cm$overall['Accuracy'])
    us_glm_precision <- c(us_glm_precision, us_cm$byClass['Precision'])
    us_glm_recall <- c(us_glm_recall, us_cm$byClass['Recall'])
    us_glm_f1 <- c(us_glm_f1, us_cm$byClass['F1'])
    
    #run the logistic regression for the oversampling
    glm_mod <- caret::train(class~.,data=os_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    os_glm_accuracy <- c(os_glm_accuracy, os_cm$overall['Accuracy'])
    os_glm_precision <- c(os_glm_precision, os_cm$byClass['Precision'])
    os_glm_recall <- c(os_glm_recall, os_cm$byClass['Recall'])
    os_glm_f1 <- c(os_glm_f1, os_cm$byClass['F1'])
    
    #run the logistic regression for the smote
    glm_mod <- caret::train(class~.,data=smote_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    smote_glm_accuracy <- c(smote_glm_accuracy, cm_smote$overall['Accuracy'])
    smote_glm_precision <- c(smote_glm_precision, cm_smote$byClass['Precision'])
    smote_glm_recall <- c(smote_glm_recall, cm_smote$byClass['Recall'])
    smote_glm_f1 <- c(smote_glm_f1, cm_smote$byClass['F1'])
    
    #run the random forest for the undersampling
    rf_mod <- caret::train(class~., data=us_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    us_rf_accuracy <- c(us_rf_accuracy, us_cm$overall['Accuracy'])
    us_rf_precision <- c(us_rf_precision, us_cm$byClass['Precision'])
    us_rf_recall <- c(us_rf_recall, us_cm$byClass['Recall'])
    us_rf_f1 <- c(us_rf_f1, us_cm$byClass['F1'])
    
    #run the random forest for the oversampling
    rf_mod <- caret::train(class~., data=os_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    os_rf_accuracy <- c(os_rf_accuracy, os_cm$overall['Accuracy'])
    os_rf_precision <- c(os_rf_precision, os_cm$byClass['Precision'])
    os_rf_recall <- c(os_rf_recall, os_cm$byClass['Recall'])
    os_rf_f1 <- c(os_rf_f1, os_cm$byClass['F1'])
    
    #run the random forest for the smote
    rf_mod <- caret::train(class~., data=smote_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    smote_rf_accuracy <- c(smote_rf_accuracy, cm_smote$overall['Accuracy'])
    smote_rf_precision <- c(smote_rf_precision, cm_smote$byClass['Precision'])
    smote_rf_recall <- c(smote_rf_recall, cm_smote$byClass['Recall'])
    smote_rf_f1 <- c(smote_rf_f1, cm_smote$byClass['F1'])
    
    #run the naive byes for the undersampling
    nb_mod <- caret::train(class~., data=us_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    us_nb_accuracy <- c(us_nb_accuracy, us_cm$overall['Accuracy'])
    us_nb_precision <- c(us_nb_precision, us_cm$byClass['Precision'])
    us_nb_recall <- c(us_nb_recall, us_cm$byClass['Recall'])
    us_nb_f1 <- c(us_nb_f1, us_cm$byClass['F1'])

    #run the naive byes for the oversampling
    nb_mod <- caret::train(class~., data=os_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    os_nb_accuracy <- c(os_nb_accuracy, os_cm$overall['Accuracy'])
    os_nb_precision <- c(os_nb_precision, os_cm$byClass['Precision'])
    os_nb_recall <- c(os_nb_recall, os_cm$byClass['Recall'])
    os_nb_f1 <- c(os_nb_f1, os_cm$byClass['F1'])
    
    #run the naive byes for the smote
    nb_mod <- caret::train(class~., data=smote_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$hypertension), mode='everything')
    smote_nb_accuracy <- c(smote_nb_accuracy, cm_smote$overall['Accuracy'])
    smote_nb_precision <- c(smote_nb_precision, cm_smote$byClass['Precision'])
    smote_nb_recall <- c(smote_nb_recall, cm_smote$byClass['Recall'])
    smote_nb_f1 <- c(smote_nb_f1, cm_smote$byClass['F1'])
  }
  
```


## Result of the hypertension analysis

The data were partitioned into a test and training set using a 70/30 split.
```{r}
df <- data.frame(us_glm_accuracy, os_glm_accuracy, smote_glm_accuracy, us_rf_accuracy, os_rf_accuracy, smote_rf_accuracy, us_nb_accuracy, os_nb_accuracy, smote_nb_accuracy)

us_glm_accuracy
os_glm_accuracy
smote_glm_accuracy
us_rf_accuracy
os_rf_accuracy
smote_rf_accuracy
us_nb_accuracy
os_nb_accuracy
smote_nb_accuracy

us_glm_precision
os_glm_precision
smote_glm_precision
us_rf_precision
os_rf_precision
smote_rf_precision
us_nb_precision
os_nb_precision
smote_nb_precision

us_glm_recall
os_glm_recall
smote_glm_recall
us_rf_recall
os_rf_recall
smote_rf_recall
us_nb_recall
os_nb_recall
smote_nb_recall

us_glm_f1
os_glm_f1
smote_glm_f1
us_rf_f1
os_rf_f1
smote_rf_f1
us_nb_f1
os_nb_f1
smote_nb_f1

c1 <- rainbow(10)
c2 <- rainbow(10, alpha=0.2)
c3 <- rainbow(10, v=0.7)
boxplot(df, col=c2, medcol=c3, whiskcol=c1, staplecol=c3, boxcol=c3, outcol=c3, pch=23, cex=2)

mean(us_nb_accuracy)
mean(us_nb_precision)
mean(us_nb_recall)
mean(us_nb_f1)

mean(os_nb_accuracy)
mean(os_nb_precision)
mean(os_nb_recall)
mean(os_nb_f1)

mean(smote_nb_accuracy)
mean(smote_nb_precision)
mean(smote_nb_recall)
mean(smote_nb_f1)

a <- matrix(
  c(mean(us_glm_accuracy),mean(us_glm_precision),mean(us_glm_recall),mean(us_glm_f1),
    mean(os_glm_accuracy),mean(os_glm_precision),mean(os_glm_recall),mean(os_glm_f1),
    mean(smote_glm_accuracy),mean(smote_glm_precision),mean(smote_glm_recall),mean(smote_glm_f1)),
  nrow=3,
  ncol=4,
  byrow = TRUE
)

a
```


## Patients with skin cancer
```{r}
#See the data before balancing
barplot(table(patients$skin_cancer), xlab=colnames(patients$skin_cancer))

#filter the dataset and have only skin_cancer disease as the target 
skin_cancer_set <- select(patients, gender, age, employment_status, education, marital_status, ancestry, available_vehicles, avg_commute,zipcode, children,daily_internet_use,military_service, skin_cancer)

#The data were partitioned into a test and training set using a 70/30 split.
train <- sample(nrow(skin_cancer_set), 0.7*nrow(skin_cancer_set), replace = FALSE)
  TrainSet <- skin_cancer_set[train,]
  ValidSet <- skin_cancer_set[-train,]
  
response <- as.factor(patients$skin_cancer)
input <- select(patients, gender, age, employment_status, education, marital_status, ancestry)
```

### Excerice the Undersampling, oversampling, and smote against the test dataset

## Logistic Regression, Randomforest, and Naive Bayes Models
```{r}
  #initialize variables
  us_glm_accuracy <- c()
  us_glm_precision <- c()
  us_glm_recall <- c()
  us_glm_f1 <- c()
  
  os_glm_accuracy <- c()
  os_glm_precision <- c()
  os_glm_recall <- c()
  os_glm_f1 <- c()
  
  smote_glm_accuracy <- c()
  smote_glm_precision <- c()
  smote_glm_recall <- c()
  smote_glm_f1 <- c()
  
  us_rf_accuracy <- c()
  us_rf_precision <- c()
  us_rf_recall <- c()
  us_rf_f1 <- c()
  
  os_rf_accuracy <- c()
  os_rf_precision <- c()
  os_rf_recall <- c()
  os_rf_f1 <- c()
  
  smote_rf_accuracy <- c()
  smote_rf_precision <- c()
  smote_rf_recall <- c()
  smote_rf_f1 <- c()
  
  us_nb_accuracy <- c()
  us_nb_precision <- c()
  us_nb_recall <- c()
  us_nb_f1 <- c()
  
  os_nb_accuracy <- c()
  os_nb_precision <- c()
  os_nb_recall <- c()
  os_nb_f1 <- c()
  
  smote_nb_accuracy <- c()
  smote_nb_precision <- c()
  smote_nb_recall <- c()
  smote_nb_f1 <- c()
  
  #use the 10-fold cross-validation and repeate the step 3 times
  train_control <- trainControl(method = "cv", number = 10)
  metric <- "Accuracy"
  mtry <- sqrt(ncol(skin_cancer_set))
  tunegrid <- expand.grid(.mtry=mtry)    
```

```{r}
  # iterate throug the sampling and model 10 times and get the mean to get the best model for the dataset prediction
  for (i in 1:10) {
    
    #run the undersampling
    data <- ubUnder(X=input, Y=response, perc=40, method="percPos")
    us_dataset <- cbind(data$X, class=data$Y)
    
    #run the oversampling
    data <- ubOver(X=input, Y=response)
    os_dataset <- cbind(data$X, class=data$Y)
    
    #run the smote
    data <- ubSMOTE(X=input, Y=response)
    smote_dataset <- cbind(data$X, class=data$Y)
    
    #use the 10-fold cross-validation and repeate the step 3 times
    train_control <- trainControl(method = "repeatedcv", number = 10, repeats=3, savePredictions = TRUE)
    
    #run the logistic regression for the undersampling
    glm_mod <- caret::train(class~.,data=us_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    us_glm_accuracy <- c(us_glm_accuracy, us_cm$overall['Accuracy'])
    us_glm_precision <- c(us_glm_precision, us_cm$byClass['Precision'])
    us_glm_recall <- c(us_glm_recall, us_cm$byClass['Recall'])
    us_glm_f1 <- c(us_glm_f1, us_cm$byClass['F1'])
    
    #run the logistic regression for the oversampling
    glm_mod <- caret::train(class~.,data=os_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    os_glm_accuracy <- c(os_glm_accuracy, os_cm$overall['Accuracy'])
    os_glm_precision <- c(os_glm_precision, os_cm$byClass['Precision'])
    os_glm_recall <- c(os_glm_recall, os_cm$byClass['Recall'])
    os_glm_f1 <- c(os_glm_f1, os_cm$byClass['F1'])
    
    #run the logistic regression for the smote
    glm_mod <- caret::train(class~.,data=smote_dataset, trControl = train_control, method="glm", family="binomial", tuneLength = 5)
    pred = predict(glm_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    smote_glm_accuracy <- c(smote_glm_accuracy, cm_smote$overall['Accuracy'])
    smote_glm_precision <- c(smote_glm_precision, cm_smote$byClass['Precision'])
    smote_glm_recall <- c(smote_glm_recall, cm_smote$byClass['Recall'])
    smote_glm_f1 <- c(smote_glm_f1, cm_smote$byClass['F1'])
    
    #run the random forest for the undersampling
    rf_mod <- caret::train(class~., data=us_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    us_rf_accuracy <- c(us_rf_accuracy, us_cm$overall['Accuracy'])
    us_rf_precision <- c(us_rf_precision, us_cm$byClass['Precision'])
    us_rf_recall <- c(us_rf_recall, us_cm$byClass['Recall'])
    us_rf_f1 <- c(us_rf_f1, us_cm$byClass['F1'])
    
    #run the random forest for the oversampling
    rf_mod <- caret::train(class~., data=os_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    os_rf_accuracy <- c(os_rf_accuracy, os_cm$overall['Accuracy'])
    os_rf_precision <- c(os_rf_precision, os_cm$byClass['Precision'])
    os_rf_recall <- c(os_rf_recall, os_cm$byClass['Recall'])
    os_rf_f1 <- c(os_rf_f1, os_cm$byClass['F1'])
    
    #run the random forest for the smote
    rf_mod <- caret::train(class~., data=smote_dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=train_control)
    pred = predict(rf_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    smote_rf_accuracy <- c(smote_rf_accuracy, cm_smote$overall['Accuracy'])
    smote_rf_precision <- c(smote_rf_precision, cm_smote$byClass['Precision'])
    smote_rf_recall <- c(smote_rf_recall, cm_smote$byClass['Recall'])
    smote_rf_f1 <- c(smote_rf_f1, cm_smote$byClass['F1'])
    
    #run the naive byes for the undersampling
    nb_mod <- caret::train(class~., data=us_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    us_cm <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    us_nb_accuracy <- c(us_nb_accuracy, us_cm$overall['Accuracy'])
    us_nb_precision <- c(us_nb_precision, us_cm$byClass['Precision'])
    us_nb_recall <- c(us_nb_recall, us_cm$byClass['Recall'])
    us_nb_f1 <- c(us_nb_f1, us_cm$byClass['F1'])

    #run the naive byes for the oversampling
    nb_mod <- caret::train(class~., data=os_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    os_cm <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    os_nb_accuracy <- c(os_nb_accuracy, os_cm$overall['Accuracy'])
    os_nb_precision <- c(os_nb_precision, os_cm$byClass['Precision'])
    os_nb_recall <- c(os_nb_recall, os_cm$byClass['Recall'])
    os_nb_f1 <- c(os_nb_f1, os_cm$byClass['F1'])
    
    #run the naive byes for the smote
    nb_mod <- caret::train(class~., data=smote_dataset, method="nb", trControl=train_control)
    pred = predict(nb_mod, newdata=ValidSet)
    cm_smote <- confusionMatrix(data=pred, as.factor(ValidSet$skin_cancer), mode='everything')
    smote_nb_accuracy <- c(smote_nb_accuracy, cm_smote$overall['Accuracy'])
    smote_nb_precision <- c(smote_nb_precision, cm_smote$byClass['Precision'])
    smote_nb_recall <- c(smote_nb_recall, cm_smote$byClass['Recall'])
    smote_nb_f1 <- c(smote_nb_f1, cm_smote$byClass['F1'])
  }
  
```


## Result of the skin cancer analysis

The data were partitioned into a test and training set using a 70/30 split.
```{r}
df <- data.frame(us_glm_accuracy, os_glm_accuracy, smote_glm_accuracy, us_rf_accuracy, os_rf_accuracy, smote_rf_accuracy, us_nb_accuracy, os_nb_accuracy, smote_nb_accuracy)

us_glm_accuracy
os_glm_accuracy
smote_glm_accuracy
us_rf_accuracy
os_rf_accuracy
smote_rf_accuracy
us_nb_accuracy
os_nb_accuracy
smote_nb_accuracy

us_glm_precision
os_glm_precision
smote_glm_precision
us_rf_precision
os_rf_precision
smote_rf_precision
us_nb_precision
os_nb_precision
smote_nb_precision

us_glm_recall
os_glm_recall
smote_glm_recall
us_rf_recall
os_rf_recall
smote_rf_recall
us_nb_recall
os_nb_recall
smote_nb_recall

us_glm_f1
os_glm_f1
smote_glm_f1
us_rf_f1
os_rf_f1
smote_rf_f1
us_nb_f1
os_nb_f1
smote_nb_f1

c1 <- rainbow(10)
c2 <- rainbow(10, alpha=0.2)
c3 <- rainbow(10, v=0.7)
boxplot(df, col=c2, medcol=c3, whiskcol=c1, staplecol=c3, boxcol=c3, outcol=c3, pch=23, cex=2)

mean(us_nb_accuracy)
mean(us_nb_precision)
mean(us_nb_recall)
mean(us_nb_f1)

mean(os_nb_accuracy)
mean(os_nb_precision)
mean(os_nb_recall)
mean(os_nb_f1)

mean(smote_nb_accuracy)
mean(smote_nb_precision)
mean(smote_nb_recall)
mean(smote_nb_f1)

a <- matrix(
  c(mean(us_glm_accuracy),mean(us_glm_precision),mean(us_glm_recall),mean(us_glm_f1),
    mean(os_glm_accuracy),mean(os_glm_precision),mean(os_glm_recall),mean(os_glm_f1),
    mean(smote_glm_accuracy),mean(smote_glm_precision),mean(smote_glm_recall),mean(smote_glm_f1)),
  nrow=3,
  ncol=4,
  byrow = TRUE
)

a
```

